### 2018-05-02 paper/abstract.tex
m

% What's the problem?
% Why is it a problem? Research gap left by other approaches?
% Why is it important? Why care?
% What's the approach? How to solve the problem?
% What's the findings? How was it evaluated, what are the results, limitations, 
% what remains to be done?
%
%abstract
%problem/topic
This paper proposes a novel fully-decentralized privacy-preserving file sharing service that eschews any third-party storage or any centralized bootstrapping service.
%research gap
Cloud-based file-sharing systems such as Dropbox require trust in the service provider and can be monitored, even when their data is encrypted. Private peer-to-peer (P2P) solutions are either particularly inefficient, or must rely on a fixed set of trusted servers for bootstrapping purposes, which limit their level of protection. % We are interested in minimizing the trust needed in other systems and preserving the users' privacy in terms of data and metadata, what they share with whom and when.
%why now
% Nowadays, many people use multiple devices and it has thus become possible to benefit from P2P technology for one's own devices alone, thereby reducing one's exposure to security and privacy threats from third parties.
%approach
In this paper, we demonstrate that an efficient and strongly private fully-decentralized data-sharing solution is possible. Our proposal combines onion routing with probabilistic node selection, informed by peer sampling and Hidden Markov Model predictions for node availability.
%findings/contributions:
Our solution benefits from sender-receiver anonymity properties provided by onion routing. %% to protect user privacy in decentralized file sharing.
We show that even a relatively naive prediction model suffices to share files despite the unreliable availability of individual nodes. %% The same model can potentially be applied for better-informed churn handling in other p2p systems. In the case of onion routing systems for anonymity, 
Our solution allows the system to include unreliable nodes and take advantage of locality, neither of which is currently provided in TOR.

>>>>

% What's the problem?
% Why is it a problem? Research gap left by other approaches?
% Why is it important? Why care?
% What's the approach? How to solve the problem?
% What's the findings? How was it evaluated, what are the results, limitations, 
% what remains to be done?
%
%abstract
%problem/topic
This paper proposes a novel fully-decentralized privacy-preserving file sharing service that eschews any third-party storage or any centralized bootstrapping service.
%research gap
Cloud-based file-sharing systems such as Dropbox require trust in the service provider and can be monitored, even when their data is encrypted. Private peer-to-peer (P2P) solutions are either particularly inefficient, or must rely on a fixed set of trusted servers for bootstrapping purposes, which limit their level of protection. % We are interested in minimizing the trust needed in other systems and preserving the users' privacy in terms of data and metadata, what they share with whom and when.
%why now
% Nowadays, many people use multiple devices and it has thus become possible to benefit from P2P technology for one's own devices alone, thereby reducing one's exposure to security and privacy threats from third parties.
%approach
In this paper, we demonstrate that an efficient and strongly private fully-decentralized data-sharing solution is possible. Our proposal combines onion routing with probabilistic node selection, informed by peer sampling and Hidden Markov Model predictions for node availability.
%findings/contributions:
Our solution benefits from sender-receiver anonymity properties provided by onion routing. %% to protect user privacy in decentralized file sharing.
We show that even a relatively naive prediction model suffices to share files despite the unreliable availability of individual nodes. %% The same model can potentially be applied for better-informed churn handling in other p2p systems. In the case of onion routing systems for anonymity, 
Our solution allows the system to include unreliable nodes, without dedicated relays, %take advantage of locality, 
neither of which is currently the case in TOR.

### 2018-05-02 paper/intro.tex
m

Although data encryption is often proposed as a first line of defense for cloud-hosted data, it is unfortunately  insufficient~\cite{granick_2017, HooffLZZ15, HarnikPS10}. Metadata such as a user'
location and activity can be still tracked, including the access
control history revealing what each user shared with whom and when. 
%  because it does not protect meta-data information
This turns out to be a critical limitation, as famously emphasized by a former NSA and CIA director: "We kill people based on metadata.''~\cite{NaughtonTheGuardian2016} 
User devices become tracking and/or monitoring \squad{}s that turn to
be the best spy over users.

@@

% approach that provides a 
and self-organizing file sharing service between the \squad{}s of different users that preserves user privacy
in terms of both data \emph{and} metadata. Our proposal hinges on two key ideas: \emph{First}, we consider proxemics relationships as a key enabler to initiate file
transfers. Close distance among users enable them to share
routing information out-of-band, e.g. from their
\squad. In contrast to state-of-the-art P2P systems, users agree in \name out-of-band to share a file, but
do not keep it available for anyone else in the future. \emph{Second}, we
rely on distributed machine learning techniques to predict the future
availability of user devices, and use this ability to implement a \emph{probabilistic predictive onion routing} mechanism. The predictive nature of our routing allows us to overcome the inherent fragility of pure P2P data-sharing schemes.

>>>>

Although data encryption is often proposed as a first line of defense for cloud-hosted data, it is unfortunately  insufficient~\cite{granick_2017, HooffLZZ15, HarnikPS10}. Metadata such as a user'
location and activity can be still tracked, including the access
control history revealing what each user shared with whom and when. 
%  because it does not protect meta-data information
This turns out to be a critical limitation, as famously emphasized by a former NSA and CIA director: "We kill people based on metadata.''~\cite{NaughtonTheGuardian2016} 
User devices become tracking and/or monitoring \squad{}s that turn out
to be the best spy over users.

@@

% approach that provides a 
and self-organizing file sharing service between the \squad{}s of different users that preserves user privacy
in terms of both data \emph{and} metadata. Our proposal hinges on two
key ideas: \emph{First},  the recent advances of smartphones and other
mobile devices enable e-squads and it has thus
become possible to benefit from P2P technology for one's own devices
alone, thereby reducing one's exposure to security and privacy threats
from third parties. Relatedly, we consider proxemics relationships as a key enabler to initiate file
transfers. Close distance among users enable them to share
routing information out-of-band, e.g. from their
\squad. In contrast to state-of-the-art P2P systems, users agree in \name out-of-band to share a file, but
do not keep it available for anyone else in the future. \emph{Second}, we
rely on distributed machine learning techniques to predict the future
availability of user devices, and use this ability to implement a
\emph{probabilistic predictive onion routing} mechanism. The
predictive nature of our routing, exploited in a stateless way, allows us to overcome the inherent
fragility of pure P2P data-sharing schemes.

### 2018-05-02 paper/message-passing-design.tex
m

To implement our  probabilistic
predictive onion routing mechanism, \name defined two different algorithms to
receive, send messages among users' \squad. For instance,
if Alice wants to send a message \(m\) to Bob, Bob uses the
\([Recv]\) algorithm (See \cref{SPORFwd}) to create a probabilistic onion route \(H_{Bob}\) from
 a set of rendez-vous point to its own devices from its \squad --- all the hops on the route are 
chosen by Bob uniformly at random from its RPS (\view).
 \([Recv]\) returns the \(H_{Bob}\).
PORs have \(L\) layers, and for each layer the number of
alternatives nodes depends of a failure threshold \(\theta\). 
As long as \(\theta\) is not exceeded, alternatives nodes are added in
\(H_{Bob}\) as depicted in \cref{ExtendRoute}. Thus the probability of
failure for the extension is \(1 - (1 - \theta)^L\). In a similar
manner, Alice create an onion route \(H_{Alice}\) to enable Bob to
acknowledge its messages. Bob and Alice exchange then their calculated
PORs (i.e. H_{Bob} andH_{Alice}), using an
out-of-band channel, \eg using one device of their \squad (See
Fig.~\ref{fig:file-exchange}~\ding{182}).

@@

Once Alice get the \(H_{Bob}\), Alice uses \([Send]\) at some
later time to send the message to Bob using any of her devices from
its \squad (See \cref{SPORFwd}). First, \([Send]\) extends the route
\(H_{Bob}\) with some hops of Alice's choosing (uniformly randomly
chosen). Second, \([Send]\) extracts from the onion the first
layer, which contained the overall
possible rendez-vous points $D$ previously selected by the recipient (i.e. Bob).
One device from $D$ is randomly chosen, and the \([Fwd]\)
sub-function is called as long as the message forwarding to one node of
the next layer is not successful (See \cref{SPORFwd}). Along the path, each node decrypts
in sequence the different layer with the current node’s private key. 
The end of the route is reached when special value \(H_{Bob} = \top\) is
encountered. Consequently the ciphered message \(c_m\) is then stored
in the disk of the local node through the \([Store]\) sub function 
(\cref{SPORFwd}).

>>>>

To implement our  probabilistic
predictive onion routing mechanism, \name defined two different algorithms to
receive, send messages among users' \squad. For instance,
if Alice wants to send a message \(m\) to Bob, Bob uses the
\([Recv]\) algorithm (See \cref{SPORFwd}) to create a probabilistic onion route \(H_{Bob}\) from
 a set of rendez-vous point to its own devices from its \squad --- all the hops on the route are 
chosen by Bob uniformly at random from its RPS (\view).
 %\([Recv]\) returns the \(H_{Bob}\).
PORs have \(L\) layers, and for each layer the number of
alternatives nodes depends of a failure threshold \(\theta\). 
As long as \(\theta\) is not exceeded, alternatives nodes are added in
\(H_{Bob}\) as depicted in \cref{ExtendRoute}. Thus the probability of
failure for the extension is \(1 - (1 - \theta)^L\). In a similar
manner, Alice create an onion route \(H_{Alice}\) to enable Bob to
acknowledge its messages. Bob and Alice exchange then their calculated
PORs (i.e. H_{Bob} and H_{Alice}), using an
out-of-band channel, \eg using one device of their \squad (See
Fig.~\ref{fig:file-exchange}~\ding{182}).

@@

Once Alice get the \(H_{Bob}\), Alice uses \([Send]\) at some
later time to send the message to Bob using any of her devices from
its \squad (See \cref{SPORFwd}). First, \([Send]\) extends the route
\(H_{Bob}\) with some hops of Alice's choosing (uniformly randomly
chosen). Second, \([Send]\) extracts from the onion the first
layer, which contained the overall
possible alternative nodes $D$.
One device from $D$ is randomly chosen, and the \([Fwd]\)
sub-function is called as long as the message forwarding to one node of
the next layer is not successful (See \cref{SPORFwd}). Along the path, each node decrypts
in sequence the different layer with the current node’s private key. 
The end of the route is reached when the special value \(H_{Bob} = \top\) is
encountered. Consequently the ciphered message \(c_m\) is then stored
in the disk of the local node through the \([Store]\) sub function 
(\cref{SPORFwd}).

In a way similar, once Bob get the \(H_{Alice}\), Bob uses \([Send]\) at some
later time to acknowledge messages sent by Alice using any of his devices from
its\squad.
